Visual proteomics using whole-lamella 2D template matching

### 2022-04-07 manuscript.tex
Minor figure [ci skip]
This build is based on
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/commit/56ce938e8d7fb4ca49b0ca90d4ceca7386ae3b0f.

This commit was created by the following CI build and job:
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/commit/56ce938e8d7fb4ca49b0ca90d4ceca7386ae3b0f/checks
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/actions/runs/2110566960

This manuscript
(\href{https://jojoelfe.github.io/deco_lace_template_matching_manuscript/v/2d0a9a5debbbf44b5f9d5797d38db68589012013/}{permalink})
was automatically generated
from \href{https://github.com/jojoelfe/deco_lace_template_matching_manuscript/tree/2d0a9a5debbbf44b5f9d5797d38db68589012013}{jojoelfe/deco\_lace\_template\_matching\_manuscript@2d0a9a5}
on March 30, 2022.

@@

We used an atomic model of the 60S mouse ribosomal subunit (6SWA) for 2D
template matching. In a subset of images the distribution of cross-correlation
scores significantly exceeded the distribution expected from non-signifcant
matching. In the resulting scaled maximum-intensity maps, clear peaks
with SNR thresholds up to 10 were apparent (Fig. \ref{fig:initmatching2}A). By
using the criterion described by for thresholding potential matches we found
that in images of cytosolic compartments we found evidence of 10-500 ribosomes in
the imaged areas. Notably we found no matches in images that were taken in the
nuclear compartment. In the cytosolic areas we found a drastically different
number of matches, In some areas we found only \textasciitilde{} 50 matches er image area, while
in another area we found more than 500 matches . This ten-fold difference in
local ribosome concentration , but we realized that current data acquistion
protocols are limited in that only a small area of the lamella is actually
imaged and, due to the manual selection of acquision positions based on the
overview image, might be biased towards cellular region that appear appealing to
the experimentalist. We therefore set out to collect cryo-EM data for 2DTM from
mammalian lamella in a high-throughput unbiased fashion.
\#\#\# DECO-LACE for 2D imaging of whole lamella

@@

We developed customized preprocessing protocol to images obtained by cryo-EMILIA
to enable their use for 2D-template matching. First we restricted calculation of
cross-correlation coefficients between individual movie frames to the central
portion of the image to prevent artifacts from the beam edges on estimation of
motion. Then we calculated a mask that defined the illuminated area of the
micrographs and used it to fill non-illumated areas with gaussian noise,
matching mean and standard deviation to the illuminated signal (Figure 3A). The
so processed images were suitable for 2D-template matching and we were able to
obtained matches with the same model used for the data in Figure 1.

@@

\begin{itemize}
\item
  Elizabeth Wright and Grant Jensen Montage tomography papers
\item
  Waffle method for higher throughput, automation of fib-milling
\item
  Throughput and bottlenecks
\item
  Visual proteomics
\item
  Granules containing ribosomes?
\item
  Threshold implications (no matches on most images)
\end{itemize}

@@

\begin{figure}
\hypertarget{fig:initmatching}{%
\centering
\includegraphics{figures/initmatching.pdf}
\caption{2D template matching of the large subunit of the ribosome in fib-milled neutrophil-like cells
(A) Overview image of the lamella. Major cellular regions are labeled, as Nucleus (Nuc), Mitochondria (M), and granular cytoplasm (GrCyt). FOVs where high-magnification images for template matching where acquired are indicated as boxes with the number of matches indicated on the bottom right. FOVs displayed in Panels B-E are color-coded.
(B-E) FOVs with projection of ribosome LSU matches shown in green. (B) Perinuclear region, the only matches are in the cytoplasmic half. (C) Cytoplasmic region with high density of ribosomes (D) Mitochondrium, as expected there are only matches in the cytoplasmic region (E) Cytoplasm, with low density of ribosomes.}\label{fig:initmatching}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:initmatching2}{%
\centering
\includegraphics{figures/initmatching2.pdf}
\caption{Maximum intensity projection cross-correlation map of micrograph shown in Figure \ref{fig:initmatching}}\label{fig:initmatching2}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:approach}{%
\centering
\includegraphics{figures/approach.pdf}
\caption{DeCo-LACE approach (A) Graphic demonstrating the data-collection strategy for DeCo-LACE}\label{fig:approach}
}
\end{figure}

\begin{figure}
\hypertarget{fig:lamella_spatial_info}{%
\centering
\includegraphics{figures/lamella_spatial_info.pdf}
\caption{This is an example-figurern}\label{fig:lamella_spatial_info}
}
\end{figure}

\begin{figure}
\hypertarget{fig:approach3}{%
\centering
\includegraphics{figures/example-figure3.pdf}
\caption{This is an example-figurern}\label{fig:approach3}
}
\end{figure}

\begin{figure}
\hypertarget{fig:approach4}{%
\centering
\includegraphics{figures/example-figure4.pdf}
\caption{This is an example-figurern}\label{fig:approach4}
}
\end{figure}

\begin{figure}
\hypertarget{fig:matching}{%
\centering
\includegraphics{figures/matching.pdf}
\caption{This is an example-figurern}\label{fig:matching}
}
\end{figure}

\begin{figure}
\hypertarget{fig:deco_lace_workflow}{%
\centering
\includegraphics{figures/deco_lace_workflow.pdf}
\caption{Workflow of DeCo-Lace processing}\label{fig:deco_lace_workflow}
}
\end{figure}

>>>>

This manuscript
(\href{https://jojoelfe.github.io/deco_lace_template_matching_manuscript/v/56ce938e8d7fb4ca49b0ca90d4ceca7386ae3b0f/}{permalink})
was automatically generated
from \href{https://github.com/jojoelfe/deco_lace_template_matching_manuscript/tree/56ce938e8d7fb4ca49b0ca90d4ceca7386ae3b0f}{jojoelfe/deco\_lace\_template\_matching\_manuscript@56ce938}
on April 7, 2022.

@@

We used an atomic model of the 60S mouse ribosomal subunit (6SWA) for 2D
template matching. In a subset of images the distribution of cross-correlation
scores significantly exceeded the distribution expected from non-signifcant
matching. In the resulting scaled maximum-intensity maps, clear peaks with SNR
thresholds up to 10 were apparent (Fig. \ref{fig:initmatching2}A). By using the
criterion described by for thresholding potential matches we found that in
images of cytosolic compartments we found evidence of 10-500 ribosomes in the
imaged areas. Notably we found no matches in images that were taken in the
nuclear compartment. In the cytosolic areas we found a drastically different
number of matches, In some areas we found only \textasciitilde{} 50 matches er image area, while
in another area we found more than 500 matches . This ten-fold difference in
local ribosome concentration , but we realized that current data acquistion
protocols are limited in that only a small area of the lamella is actually
imaged and, due to the manual selection of acquision positions based on the
overview image, might be biased towards cellular region that appear appealing to
the experimentalist. We therefore set out to collect cryo-EM data for 2DTM from
mammalian lamella in a high-throughput unbiased fashion.

\hypertarget{deco-lace-for-2d-imaging-of-whole-lamella}{%
\subsubsection{DECO-LACE for 2D imaging of whole lamella}\label{deco-lace-for-2d-imaging-of-whole-lamella}}

@@

We developed customized preprocessing protocol to images obtained by cryo-EMILIA
to enable their use for 2D-template matching. First we restricted calculation of
cross-correlation coefficients between individual movie frames to the central
portion of the image to prevent artifacts from the beam edges on estimation of
motion {[}Figure S{]}. Then we calculated a mask that defined the illuminated area
of the micrographs and used it to fill non-illumated areas with gaussian noise,
matching mean and standard deviation to the illuminated signal (Figure S). The
so processed images were suitable for 2D-template matching and we were able to
obtained matches with the same model used for the data in Figure 1.

We then matched all individual tiles for ribosomes and assembled a montaged view
as described above (Figure \ref{fig:matching} A+G). The resulting montages reveal
ribsome distribution throughout the FIB-milled slice of individual cells.
Organelles like the nucleus and mitochondria only show sporadic matches with low
SNRs, consistent with the estimated false-positive rate on one per tile. During
the assembly we calculated the Z positions from the indivdual estimated defocus
levels and the focus plane of the microscope during acquisition of each tile.
When viewed from the side the ribosome position therefore show the slope of the
lamella in the microscope frame. Furthermore, the positions are also consistent
with an uneven thickness of the lamella, thinner at the leading edge, even
though, lamella were generated using the ``overtilt'' approach {[}cite{]}.

@@

\begin{itemize}
\item
  Elizabeth Wright and Grant Jensen Montage tomography papers
\item
  Waffle method for higher throughput, automation of fib-milling
\item
  Throughput and bottlenecks
\item
  Visual proteomics
\item
  Granules containing ribosomes?
\item
  Threshold implications (no matches on most images)
\item
  K3 not linear
\end{itemize}

@@

\begin{figure}
\hypertarget{fig:initmatching}{%
\centering
\includegraphics{figures/initmatching.png}
\caption{2D template matching of the large subunit of the ribosome in fib-milled neutrophil-like cells
(A) Overview image of the lamella. Major cellular regions are labeled, as Nucleus (Nuc), Mitochondria (M), and granular cytoplasm (GrCyt). FOVs where high-magnification images for template matching where acquired are indicated as boxes with the number of matches indicated on the bottom right. FOVs displayed in Panels B-E are color-coded.
(B-E) FOVs with projection of ribosome LSU matches shown in green. (B) Perinuclear region, the only matches are in the cytoplasmic half. (C) Cytoplasmic region with high density of ribosomes (D) Mitochondrium, as expected there are only matches in the cytoplasmic region (E) Cytoplasm, with low density of ribosomes.}\label{fig:initmatching}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:initmatching2}{%
\centering
\includegraphics{figures/initmatching2.png}
\caption{2D template matching of the large subunit of the ribosome in fib-milled
neutrophil-like cells (A) Maximum intensity projection cross-correlation map of
micrograph shown in Figure
\ref{fig:initmatching} (B+C) 3D plot of MIP regions indicated by color boxes in Panel A}\label{fig:initmatching2}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:approach}{%
\centering
\includegraphics{figures/approach.png}
\caption{DeCo-LACE approach (A) Graphic demonstrating the data-collection strategy for
DeCo-LACE. The electron beam is condensed to a diameter \(D_{Beam}\) that allows captured of
the whole illuminated area on the camera. Beam-image shift along X and Y
(\(BIS_X\),\$\(BIS_Y\)) is used to raster the whole lamella
(B) Diagram of the collection algorithm
(C) Example overview image of a lamella with the designated acquisition
positions and the used beam diameter indicated with red circles
(D+E) Representative micrographs takne with a condensed beam at eucentric focus
(D) or fringe-free focus (E)
(F) Boxplot of defocus measured by ctffind of micrographs taken by the DeCo-Lace
approach on 4 lamella images at eucentric focus and 4 lamella imaged with
fringe-free focus.
(F+G) Lamella overview images of lamella imaged at eucentric focus (F) Overview
image taken at low magnification (40Å pixel size) (G) Overview created by
montaging high magnification images taken with the DeCo-Lace approach (1.5Å
pixelsize)
(H+I) Lamella overview images of lamella imaged at fringe-free focus (H) Overview
image taken at low magnification (40Å pixel size) (I) Overview created by
montaging high magnification images taken with the DeCo-Lace approach (1.5Å
pixelsize)}\label{fig:approach}
}
\end{figure}

\begin{fignos:tagged-figure}[S2]

\begin{figure}
\hypertarget{fig:lamella_images}{%
\centering
\includegraphics{figures/lamella_images.png}
\caption{Overview images of lamellae imaged using the DeCo-LACE approach taken at low-magnification}\label{fig:lamella_images}
}
\end{figure}

\begin{fignos:tagged-figure}[S3]

\begin{figure}
\hypertarget{fig:lamella_spatial_info}{%
\centering
\includegraphics{figures/lamella_spatial_info.png}
\caption{Defocus estimation of individual tiles of DeCo-Lace montages
(A) Defocus values of individual micrographs taken using the DeCo-Lace approach
plotted as a function of the Beam-Image-Shift values.
(B) Defocus astigmatism of individual micrographs taken using the DeCo-Lace approach
plotted as a function of the Beam-Image-Shift values.}\label{fig:lamella_spatial_info}
}
\end{figure}

\begin{fignos:tagged-figure}[S4]

\begin{figure}
\hypertarget{fig:crop_unblur}{%
\centering
\includegraphics{figures/crop_unblur.png}
\caption{Motion correction of movies with condensed beams.
At the top of each panel is an average of the movie that was motion-corrected
with a red dashed box indicating the region that was used to estimate shifts.
Below is a graph indicating the estimated shifts of the individual frames of the
movie. Below this is the MIP of 2DTM using the large subunit of the mouse ribosome.
(A) Motion correction of the whole movie
(B) Notion correction of a cropped region of the movie that eliminates the beam
edges
(C) Motion correction of the whole movie, using only hte cropped region to
estimate the shifts}\label{fig:crop_unblur}
}
\end{figure}

\begin{fignos:tagged-figure}[S5]

\begin{figure}
\hypertarget{fig:lamella_motion_thickness}{%
\centering
\includegraphics{figures/lamella_motion_thickness.png}
\caption{Motion correction of individual tiles imaged using the DeCo-LACE approach
(A) Total estimated motion of individual micrographs taken using the DeCo-Lace approach
plotted as a function of the Beam-Image-Shift values.
(B) Electron intensity of individual micrographs taken using the DeCo-Lace approach
plotted as a function of the Beam-Image-Shift values.}\label{fig:lamella_motion_thickness}
}
\end{figure}

\begin{figure}
\hypertarget{fig:matching_stat}{%
\centering
\includegraphics{figures/matching_stat.png}
\caption{This is an example-figurern}\label{fig:matching_stat}
}
\end{figure}

\begin{figure}
\hypertarget{fig:matching_euc}{%
\centering
\includegraphics{figures/matching_euc.png}
\caption{Template matching in lamella imaged using the DeCo-Lace approach at eucentric focus}\label{fig:matching_euc}
}
\end{figure}

\begin{figure}
\hypertarget{fig:matching_fff}{%
\centering
\includegraphics{figures/matching_fff.png}
\caption{This is an example-figurern}\label{fig:matching_fff}
}
\end{figure}

\begin{figure}
\hypertarget{fig:deco_lace_workflow}{%
\centering
\includegraphics{figures/deco_lace_workflow.png}
\caption{Workflow of DeCo-Lace processing}\label{fig:deco_lace_workflow}
}
\end{figure}
